---
title: "compare-nearest-taz"
output: html_document
---

```{r}
library(tidyverse)
library(sf)
library(mapview)
library(foreign)
```

# Introduction
In this markdown, we compare the outputs of the TagLinksWithTAZID function in the 01_Update_Link_Node_TAZID python script completed with arcpy and with geopandas libraries. For some reason, the nearest function finds a different set of "closest" TAZID's to the link when done with arcpy opposed to geopandas. We want to validate our results, and understand why this is happening.

# Read in and Organize Data
```{r}
link_taz_sj_arcpy <- read.dbf("compare/Link_TAZ_SJ_deleteTemp-arcpy.dbf") %>%
  select(A,B,X_MID,Y_MID,TAZID_1,TAZID_V832) %>%
  rename("TAZID_1_arcpy" = "TAZID_1", "TAZID_V832_arcpy" = "TAZID_V832", "X_MID_arcpy" = "X_MID", "Y_MID_arcpy" = "Y_MID")
link_taz_sj_geo <-   read.dbf("compare/Link_TAZ_SJ_deleteTemp-geo.dbf") %>%
  #read.dbf("temp_gpd_geo1/Link_TAZ_SJ_deleteTemp.dbf") %>% 
  select(A,B,X_MID,Y_MID,TAZID_1,TAZID_V832) %>%
  rename("TAZID_1_geo" = "TAZID_1", "TAZID_V832_geo" = "TAZID_V832", "X_MID_geo" = "X_MID", "Y_MID_geo" = "Y_MID")
```

Now that we have read in the two files, lets join them together and look at the attributes for where the tazids don't line up. The following code shows that for all except two links, the x or y value doesn't match up between arcpy and geopands. This could be a leading factor for the discrepancy. For the two links that do have the same x and y value, the links exist exactly at the boundary of two tazes.

```{r}
link_taz_sj_comp <- link_taz_sj_arcpy %>%
  left_join(link_taz_sj_geo, by = c("A","B")) %>%
  mutate(diff = ifelse(TAZID_1_arcpy == TAZID_1_geo, 1,0),
         xdiff = ifelse(X_MID_arcpy == X_MID_geo, 1,0),
         ydiff = ifelse(Y_MID_arcpy == Y_MID_geo, 1, 0)) %>%
  filter(diff == 0)
```

# Visualize the questionable locations
```{r}
c1_link <- st_read("inputs-04-14/C1_Link.shp") %>%
  select(A,B)
taz <- st_read("inputs-04-14/TAZ.shp")
```

```{r}
sf_link_taz_sj_comp <- link_taz_sj_comp %>%
  left_join(c1_link, by = c("A","B")) %>%
  st_as_sf()
sf_link_taz_sj_comp_centroid <- sf_link_taz_sj_comp %>% 
  st_as_sf() %>%
  st_point_on_surface()
```


```{r}
mapview(taz) + 
  mapview(sf_link_taz_sj_comp) +
  mapview(sf_link_taz_sj_comp_centroid)
```
The map shows that all the questionable links exist along a boundary of a taz location. This means that it probably doesn't really matter which taz is chosen. However, lets dive a little deeper.

```{r}
link_taz_sj_comp_arcpy <- link_taz_sj_comp %>%
  select(A,B,X_MID_arcpy,Y_MID_arcpy) %>%
  st_as_sf(coords=c(3:4)) %>%
  st_set_crs(26912) %>% 
  st_transform(4326)
link_taz_sj_comp_geo <- link_taz_sj_comp %>%
  select(A,B,X_MID_geo,Y_MID_geo) %>%
  st_as_sf(coords=c(3:4)) %>%
  st_set_crs(26912) %>% 
  st_transform(4326)
  
```

This map verifies that the differences in the precision of the coordinate is NOT the reason why the geopandas calculation is different from the arcpy calculation. 

```{r}
mapview(list(taz,
             sf_link_taz_sj_comp,
             sf_link_taz_sj_comp_centroid,
             link_taz_sj_comp_arcpy,
             link_taz_sj_comp_geo), 
        col.regions=list(
              "gray",
              "blue",
              "blue",
              "red",
              "green"), 
        col = list(
              "black",
              "blue", 
              "blue",
              "red",
              "green"))
```

I'm actually confused now because geopandas was supposed to find which taz each link midpoint exists in. It does this with the within function. BUT I THINK that for these links shown above, the "within" join returned NA. I need to find the limitation with the "within" function here, because it should be getting the taz that exists. The nearest function that I wrote should only be used for links that exist outside the taz boundaries.

Next Steps to figure out:

- Are there links that exist outside the taz boundaries (if not, do I need the nearest function at all?)
  - YES - so the nearest function is needed (see maps below)
  
```{r}
midpointsNotFound <- read_csv('compare/midpointsNotFound.csv') %>%
 st_as_sf(coords=c(6:7)) %>%
  st_set_crs(26912) %>% 
  st_transform(4326)
mapview(list(taz,
             sf_link_taz_sj_comp,
             midpointsNotFound))
```

```{r}
midpointsNotFound2 <- read_csv('compare/midpointsNotFound2.csv') %>%
  filter(is.na(TAZID_1))%>%
 st_as_sf(coords=c(4:5)) %>%
  st_set_crs(26912) %>% 
  st_transform(4326)
mapview(list(taz,
             sf_link_taz_sj_comp,
             midpointsNotFound2))
             
```

  
  - I also figured out that the nearest function and the "within" function are calculating the same value. This means that it is a deeper problem than it seems.
  

# Conclusion

The results of this test are technically inconclusive, but the following findings will help us move forward

- The nearest function is necessary for about 45 of the link midpoints that fall outside the taz polygon boundaries
- Geopanda's "within" sjoin operation does not match up with arcpy's within operation for about 40 occurrences. All 40 of these occurrences fall along the border of two taz polygons. 
- **Since all mismatches occur along the boundary, we are not going to further investigate why geopandas is assigning a different polygon than arcpy -- it doesn't really matter if it is along the boundary**

This Rmarkdown makes it look like that geopandas isn't as precise as arcpy, as when you zoom into the mismatched points, it seems that arcpy gets it correct. We cannot be sure of this conclusion, however, because what if R and arcpy are actually less precise whereas geopandas is just more precise than we can view? Overall, we do not know why geopandas isn't matching up with arcpy, however, we decide that it doesn't matter because the locations the mismatches occur could go either way.
